---
title: "example_analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{example_analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

rm(list = ls())
```

Load the package

```{r setup}
library(systemsseRology)
```

Read the data from an Excel file. Z-score the data and assign the label in the variable y.

```{r data}
library(readxl)
data <- as.data.frame(read_excel("../data/RSTR_data_NatureMedicine_subset.xlsx"))

X <- as.matrix(as.data.frame(lapply(data[, 5:ncol(data)], as.numeric)))
X_zscore <- scale(X, center = TRUE, scale = TRUE)
y <- factor(data$y, levels = unique(data$y))
```

The variable df_features contains information about the features, mainly used for visualization routines. 
Here, we assign a label (removing underscores) and extract the antigen. 

```{r annotation}
df_features <- data.frame(name = colnames(X))
df_features$label <- gsub("_", " ", df_features$name)
df_features$antigen <- gsub("_.*", "", df_features$name)
df_features$antigen <- factor(df_features$antigen)
df_features$feature_class <- rep(NA, length = nrow(df_features))
df_features$feature_class[which(grepl("IgG|IgA|IgM", df_features$name))] <- "titer"
df_features$feature_class[which(grepl("FcR", df_features$name))] <- "FcR"
df_features$feature_class <- factor(df_features$feature_class)
```

Example PCA and PLS-DA (using the package ropls) and visualization.

```{r PLSDA vis}
library(RColorBrewer)

# Define colors
my_colors = list(group = c(RSTR = "red", LTB = "blue"))
my_colors$antigen <- colorRampPalette(brewer.pal(name = "Dark2", n = 8))(nlevels(df_features$antigen))
names(my_colors$antigen) <- unique(df_features$antigen)
my_colors$feature_class <- colorRampPalette(brewer.pal(name = "Set1", n = 3))(nlevels(df_features$feature_class))
names(my_colors$feature_class) <- unique(df_features$feature_class)

# general options for plotting
opts_plot <- list(df_features = df_features,
                  loading_alpha = 1, # transparency for the loadings
                  score_alpha = 1, # transparency for the scores
                  LV_ind = c(1,2), # which LVs to plot
                  color_features = "antigen", # according to which property (defined in df_features) the features should be color-coded
                  colors = my_colors) 

# Perform a simple PCA using the interface function pca_ropls
model_pca <- pca_ropls(X)
plt_scores_pca <- visualize_ropls_scores(model_pca, y, options = opts_plot)
print(plt_scores_pca)

# Perform a PLS-DA and plot the scores and loadings
model <- train_ropls(X, y)

plt_scores <- visualize_ropls_scores(model, y, options = opts_plot)
print(plt_scores)

plt_loadings <- visualize_ropls_loadings(model, options = opts_plot)
print(plt_loadings)

# set additional options required to color code enrichment in the bar plot of the loadings
opts_plot$X <- X
opts_plot$y <- y
opts_plot$LV_ind <- 2
opts_plot$mark_enrichment <- TRUE
plt_loadings_bar <- visualize_ropls_loadings_bar(model, options = opts_plot)
print(plt_loadings_bar)
```

Analyze model results. First we check the accuracy obtained on the whole data set. 

```{r cross-validate pls}
y_pred <- predict_ropls(model, X)
acc <- score_accuracy(y, y_pred)

print(paste("Performance on full data set:", acc, "accuracy"))

method = list(select = select_lasso,
              train = train_ropls,
              predict = predict_ropls,
              score = score_accuracy)

opts = list(n_folds = 5, rf_trials = 0, pt_trials = 0)

return_vals <- validate(X, y, method, opts)

print(paste("Performance in 5-fold cv:", return_vals$cv_score, "accuracy"))
```

Permutation testing to assess model significance using two different approaches.
1) Size-matched random features
2) Shuffled labels

```{r model validation}
opts$rf_trials <- 10
opts$pt_trials <- 10
return_vals <- validate(X, y, method, opts)
```

